Commutative Character Sheaves

### 2015-11-12 followup.tex
Added a proof of a lemma for Section 3.

\begin{theorem}\label{thm:Gab}
Pull-back along the abelianization $q : G \to G\ab$
\[
q^* : \CCS(G\ab) \to \CCS(G)
\]
defines an equivalence of categories.
\end{theorem}

@@

Let $G$ be as above.
Let $H$ be a group scheme and write $n : H \times H \to H$ for the multiplication morphism.
Let $a : H \times G\to G$ be a group action compatible with the group structure on $G$ and write $p : H\times G \to G$ for projection. 
Consider the morphisms
\[
\begin{tikzcd}
H \times H \times G 
\arrow[shift left=4]{r}{b_1, b_2, b_3} 
\arrow{r}{}
\arrow[shift right=4]{r}{} 
& H \times G 
\arrow[shift left=2]{r}{a}
\arrow[shift right=2]{r}[swap]{p}
 & G 
\end{tikzcd}
\]
defined by $b_1(h_1,h_2,g) = (n(h_1,h_2),g)$,  $b_2(h_1,h_2,g) = (h_1,a(h_2,g))$ and $b_3(h_1,h_2,g) = (h_2,g)$.
Then, $a\circ b_1 = a\circ b_2$ and $a\circ b_3 = p\circ b_2$ and $p\circ b_1 = p\circ b_3$.
An $H$-equivariant Weil local sytem on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  
\[
\nu : a^*\cs{L} \to p^*\cs{L}
\] 
of Weil local systems on $H\times G$ such that the following diagram of isomorphisms of local systems on $H\times H\times G$ commutes,
\[
\begin{tikzcd}
\ &  \arrow{dl}{a\circ b_1 = a\circ b_2}  b_2^*  a^*\, \mathcal{L} \arrow{rr}{b_2^*(\nu)} && b_2^* p^*\, \mathcal{L} \arrow{dr}[swap]{p\circ b_2 = a\circ b_3} & \\
b_1^* a^*\, \mathcal{L} \arrow{dr}{b_1^*(\nu)} &&&&   \arrow{dl}[swap]{b_3^*(\nu)} b_3^* a^*\, \mathcal{L} \\
& b_1^* p^*\, \mathcal{L}  && \arrow{ll}[swap]{p\circ b_3 = p\circ b_1} b_3^* p^*\, \mathcal{L}, & 
\end{tikzcd}
\]
and $s^*(\nu) = \id_{\cs{L}}$.
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations above; so, for example, the top left isomorphism comes from $b_1^* a^* \iso b_2^* a^*$ since $a\circ b_1 = a\circ b_2$.
\iffalse
On stalks, this condition is ...
\[
\begin{tikzcd}
\ &  \cs{L}_{h_1\cdot(h_2\cdot g)} \arrow{r}{\nu_{h_1,h_2\cdot g}} & \cs{L}_{h_2\cdot g} & \\
\cs{L}_{(h_1 h_2) \cdot g} \arrow{dr}[swap]{\nu_{(h_1h_2),g}} \arrow[equal]{ur} &&& \arrow[equal]{ul}  \arrow{dl}{\nu_{h_2,g}} \cs{L}_{h_2\cdot g} \\
& \cs{L}_g \arrow[equal]{r} & \cs{L}_g & 
\end{tikzcd}
\]
\fi
Morphisms of $H$-equivariant Weil local systems $(\cs{L},\nu)\to (\cs{L}',\nu')$ are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which the diagram
\[
\begin{tikzcd}
\arrow{d}[swap]{\nu} a^*\cs{L} \arrow{r}{a^*(\alpha)} & a^*\cs{L}' \arrow{d}{\nu'} \\
p^*\cs{L} \arrow{r}{p^*(\alpha)} & p^*\cs{L}'
\end{tikzcd}
\]
commutes.
This defines $\Loc_H(G)$, the category of $H$-equivariant Weil local systems on $G$.

@@

An $H$-equivariant linear character sheaf on $G$ is a triple $(\cs{L},\mu, \nu)$ where $(\cs{L},\mu)$ is a linear character sheaf and $\nu$ is compatible with $\mu$ in the sense. 
Consider the following morphisms: $c_0 : H\times G \times G\to H\times G \times H \times G$ defined by $c_0(h,g_1,g_2)\ceq(h,g_1,h,g_2)$;
$c_1 : H \times G\times G \to G\times G$ defined by $c_1(h, g_1, g_2) \ceq (a(h,g_1),a(h,g_2))$;  $c_2 : H \times G\times G \to H\times G$ defined by $c_2(h,g_1,g_2) \ceq (h, m(g_1,g_2))$; and $c_3 : H \times G\times G \to G\times G$ defined by $c_3(h, g_1, g_2) = (g_1, g_2)$. 
Then $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ if $(\cs{L},\mu)$ is a linear character sheaf on $G$ and the following diagram of Weil local systems on $H \times G\times G$ commutes.
\[
\begin{tikzcd}
\ & \arrow{dl}{c_2\circ a = m\circ c_1}  c_2^* a^* \cs{L} \arrow{r}{c_2^*(\nu)} & c_2^* p^* \cs{L} \arrow{dr}[swap]{p\circ c_2 = m\circ c_3}  &  \\
\arrow{d}[swap]{c_1^*(\mu)} c_1^* m^* \cs{L} &&& c_3^* m^*\cs{L} \arrow{d}{c_3^*(\mu)} \\
c_1^*(\cs{L}\boxtimes \cs{L}) \arrow{dr}{p_i\circ c_1 = a\circ p_i\circ c_0} &&& \arrow{dl}[swap]{p\circ p_i\circ c_0 = p_i \circ c_3} c_3^*(\cs{L}\boxtimes \cs{L}) \\
 & c_0^*(a^*\cs{L}\boxtimes a^*\cs{L}) \arrow{r}{c_0^*(\nu\boxtimes \nu)} & c_0^*(p^*\cs{L} \boxtimes p^*\cs{L}) & 
\end{tikzcd}
\]
\iffalse
On stalks, this condition is
\[
\begin{tikzcd}
\arrow{d}[swap]{\mu_{(h\cdot g_1, h\cdot g_2)}} \cs{L}_{(h\cdot g_1)(h\cdot g_2)} & \arrow[equal]{l}  \cs{L}_{h\cdot(g_1 g_2)} \arrow{r}{\nu_{(h, g_2 g_2)}} & \cs{L}_{g_1 g_2} \arrow[equal]{r} & \cs{L}_{g_1 g_2} \arrow{d}{\mu_{(g_1,g_2)}} \\
\cs{L}_{h\cdot g_1}\otimes \cs{L}_{h\cdot g_2} \arrow[equal]{r} & \cs{L}_{h\cdot g_1}\otimes \cs{L}_{h\cdot g_2} \arrow{r}{\nu_{(h,g_1)}\otimes \nu_{(h,g_2)}} & \cs{L}_{g_1}\otimes \cs{L}_{g_2} & \arrow[equal]{l} \cs{L}_{g_1} \otimes \cs{L}_{g_2},
\end{tikzcd}
\]
for all $(h,g_1,g_2)\in H\times G\times G$.
\fi

A morphism of $H$-equivariant linear character sheaves $(\cs{L},\mu,\nu) \to (\cs{L}',\mu',\nu')$ is a morphism of $H$-equivariant Weil sheaves $\alpha : \cs{L}\to \cs{L}'$ which is also a morphism of linear character sheaves, so
\[
\begin{tikzcd}
\arrow{d}[swap]{\mu} m^*\cs{L} \arrow{r}{m^*(\alpha)} & m^*\cs{L}' \arrow{d}{\mu'} \\
\cs{L}\boxtimes \cs{L} \arrow{r}{\alpha\boxtimes\alpha} & \cs{L}'\boxtimes\cs{L}' 
\end{tikzcd}
\]
commutes.
%
Let $\CS_H(G)$ be the category of $H$-equivariant linear character sheaves on $G$.

@@

>>>>

\begin{theorem}\label{thm:Gab}
Pull-back along the abelianization $q : G \to G\ab$
defines an equivalence of categories
\[
\CCS(G\ab) \to \CCS(G).
\]
\end{theorem}

@@

Let $G$ be as above.
Let $H$ be a group scheme and write $n : H \times H \to H$ for the multiplication morphism.
Let $a : H \times G\to G$ be a group action compatible with the group structure on $G$ and write $p : H\times G \to G$ for projection. 
Consider the morphisms
\[
\begin{tikzcd}
H \times H \times G 
\arrow[shift left=4]{r}{b_1, b_2, b_3} 
\arrow{r}{}
\arrow[shift right=4]{r}{} 
& H \times G 
\arrow[shift left=2]{r}{a}
\arrow[shift right=2]{r}[swap]{p}
 & G 
\end{tikzcd}
\]
defined by $b_1(h_1,h_2,g) = (n(h_1,h_2),g)$,  $b_2(h_1,h_2,g) = (h_1,a(h_2,g))$ and $b_3(h_1,h_2,g) = (h_2,g)$.
Then, $a\circ b_1 = a\circ b_2$ and $a\circ b_3 = p\circ b_2$ and $p\circ b_1 = p\circ b_3$.
An $H$-equivariant Weil local sytem on $G$ is a Weil local system $\cs{L}$ on $G$ together with an isomorphism  
\[
\nu : a^*\cs{L} \to p^*\cs{L}
\] 
of Weil local systems on $H\times G$ such that the following diagram of isomorphisms of local systems on $H\times H\times G$ commutes,
\begin{equation}\label{eq:nu}
\begin{tikzcd}
\ &  \arrow{dl}{a\circ b_1 = a\circ b_2}  b_2^*  a^*\, \mathcal{L} \arrow{rr}{b_2^*(\nu)} && b_2^* p^*\, \mathcal{L} \arrow{dr}[swap]{p\circ b_2 = a\circ b_3} & \\
b_1^* a^*\, \mathcal{L} \arrow{dr}{b_1^*(\nu)} &&&&   \arrow{dl}[swap]{b_3^*(\nu)} b_3^* a^*\, \mathcal{L} \\
& b_1^* p^*\, \mathcal{L}  && \arrow{ll}[swap]{p\circ b_3 = p\circ b_1} b_3^* p^*\, \mathcal{L}, & 
\end{tikzcd}
\end{equation}
and $s^*(\nu) = \id_{\cs{L}}$.
In the diagram above, the arrows labeled with equations come from canonical isomorphisms of functors on Weil sheaves derived from the equations above; so, for example, the top left isomorphism comes from $b_1^* a^* \iso b_2^* a^*$ since $a\circ b_1 = a\circ b_2$.
\iffalse
On stalks, this condition is ...
\[
\begin{tikzcd}
\ &  \cs{L}_{h_1\cdot(h_2\cdot g)} \arrow{r}{\nu_{h_1,h_2\cdot g}} & \cs{L}_{h_2\cdot g} & \\
\cs{L}_{(h_1 h_2) \cdot g} \arrow{dr}[swap]{\nu_{(h_1h_2),g}} \arrow[equal]{ur} &&& \arrow[equal]{ul}  \arrow{dl}{\nu_{h_2,g}} \cs{L}_{h_2\cdot g} \\
& \cs{L}_g \arrow[equal]{r} & \cs{L}_g & 
\end{tikzcd}
\]
\fi
Morphisms of $H$-equivariant Weil local systems $(\cs{L},\nu)\to (\cs{L}',\nu')$ are morphisms of Weil local systems $\alpha: \cs{L}\to \cs{L}'$ for which the diagram
\begin{equation}\label{eq:alphanu}
\begin{tikzcd}
\arrow{d}[swap]{\nu} a^*\cs{L} \arrow{r}{a^*(\alpha)} & a^*\cs{L}' \arrow{d}{\nu'} \\
p^*\cs{L} \arrow{r}{p^*(\alpha)} & p^*\cs{L}'
\end{tikzcd}
\end{equation}
commutes.
This defines $\Loc_H(G)$, the category of $H$-equivariant Weil local systems on $G$.

@@

An $H$-equivariant linear character sheaf on $G$ is a triple $(\cs{L},\mu, \nu)$ where $(\cs{L},\mu)$ is a linear character sheaf and $\nu$ is compatible with $\mu$ in the sense. 
Consider the following morphisms: $c_0 : H\times G \times G\to H\times G \times H \times G$ defined by $c_0(h,g_1,g_2)\ceq(h,g_1,h,g_2)$;
$c_1 : H \times G\times G \to G\times G$ defined by $c_1(h, g_1, g_2) \ceq (a(h,g_1),a(h,g_2))$;  $c_2 : H \times G\times G \to H\times G$ defined by $c_2(h,g_1,g_2) \ceq (h, m(g_1,g_2))$; and $c_3 : H \times G\times G \to G\times G$ defined by $c_3(h, g_1, g_2) = (g_1, g_2)$. 
Then $(\cs{L},\mu,\nu)$ is an $H$-equivariant linear character sheaf on $G$ if $(\cs{L},\mu)$ is a linear character sheaf on $G$ and the following diagram of Weil local systems on $H \times G\times G$ commutes.
\begin{equation}\label{eq:munu}
\begin{tikzcd}
\ & \arrow{dl}{c_2\circ a = m\circ c_1}  c_2^* a^* \cs{L} \arrow{r}{c_2^*(\nu)} & c_2^* p^* \cs{L} \arrow{dr}[swap]{p\circ c_2 = m\circ c_3}  &  \\
\arrow{d}[swap]{c_1^*(\mu)} c_1^* m^* \cs{L} &&& c_3^* m^*\cs{L} \arrow{d}{c_3^*(\mu)} \\
c_1^*(\cs{L}\boxtimes \cs{L}) \arrow{dr}{p_i\circ c_1 = a\circ p_i\circ c_0} &&& \arrow{dl}[swap]{p\circ p_i\circ c_0 = p_i \circ c_3} c_3^*(\cs{L}\boxtimes \cs{L}) \\
 & c_0^*(a^*\cs{L}\boxtimes a^*\cs{L}) \arrow{r}{c_0^*(\nu\boxtimes \nu)} & c_0^*(p^*\cs{L} \boxtimes p^*\cs{L}) & 
\end{tikzcd}
\end{equation}
\iffalse
On stalks, this condition is
\[
\begin{tikzcd}
\arrow{d}[swap]{\mu_{(h\cdot g_1, h\cdot g_2)}} \cs{L}_{(h\cdot g_1)(h\cdot g_2)} & \arrow[equal]{l}  \cs{L}_{h\cdot(g_1 g_2)} \arrow{r}{\nu_{(h, g_2 g_2)}} & \cs{L}_{g_1 g_2} \arrow[equal]{r} & \cs{L}_{g_1 g_2} \arrow{d}{\mu_{(g_1,g_2)}} \\
\cs{L}_{h\cdot g_1}\otimes \cs{L}_{h\cdot g_2} \arrow[equal]{r} & \cs{L}_{h\cdot g_1}\otimes \cs{L}_{h\cdot g_2} \arrow{r}{\nu_{(h,g_1)}\otimes \nu_{(h,g_2)}} & \cs{L}_{g_1}\otimes \cs{L}_{g_2} & \arrow[equal]{l} \cs{L}_{g_1} \otimes \cs{L}_{g_2},
\end{tikzcd}
\]
for all $(h,g_1,g_2)\in H\times G\times G$.
\fi

A morphism of $H$-equivariant linear character sheaves $(\cs{L},\mu,\nu) \to (\cs{L}',\mu',\nu')$ is a morphism of $H$-equivariant Weil sheaves $\alpha : \cs{L}\to \cs{L}'$ which is also a morphism of linear character sheaves, so
\begin{equation}\label{eq:alphanu}
\begin{tikzcd}
\arrow{d}[swap]{\mu} m^*\cs{L} \arrow{r}{m^*(\alpha)} & m^*\cs{L}' \arrow{d}{\mu'} \\
\cs{L}\boxtimes \cs{L} \arrow{r}{\alpha\boxtimes\alpha} & \cs{L}'\boxtimes\cs{L}' 
\end{tikzcd}
\end{equation}
commutes.
%
Let $\CS_H(G)$ be the category of $H$-equivariant linear character sheaves on $G$.

@@

\begin{proof}
Pulling back $\cs{L}$ along $m : G\times G \to G$ produces the linear character sheaf $(m^*\cs{L}, m_2^*(\mu))$, where $m_2 : G\times G\times G\times G\to G\times G$ is defined by $m_2(x_1,y_1,x_2,y_2) = (m(x_1,x_2), m(y_1,y_2)$;
this can be seen by following the argument of [Dictionary].
Set $\cs{L}_2\ceq m^*\cs{L}$ and $\mu_2 \ceq m_2^*(\mu)$.
The linear character sheaf $(\cs{L}_2,\mu_2)$ comes equipped with the isomorphism $\nu_2 : a_2^*\cs{L}_2\to p_2^*\cs{L}_2$ where $a_2 : H\times H\times G\times G\to G\times G$ is given by $a_2(h_1,g_1,h_2,g_2) = (a(h_1,g_1), a(h_2,g_2)$ and $p_2 : H\times H\times G\times G\to G\times G$ is given by $p_2(h_1,g_1,h_2,g_2)= ( g_1,g_2)$.  
This isomorphism satisfies \eqref{eq:nu} and \eqref{eq:munu}.
Likewise $\cs{L}\boxtimes \cs{L}$ is an $H\times H$-equivariant linear character sheaf with isomorphism $\nu \boxtimes\nu$, as indicated in the diagram below.
\[
\begin{tikzcd}
\arrow{d}[swap]{\nu_2} a_2^*(m^*\cs{L}) \arrow{r}{a_2^*(\mu)} & a_2^*(\cs{L}\boxtimes\cs{L})  \arrow{r} & (a^*\cs{L})\boxtimes (a^*\cs{L}) \arrow{d}{\nu\boxtimes\nu} \\
p_2^*(m^*\cs{L}) \arrow{r}{p_2^*(\mu)} & p_2^*(\cs{L}\boxtimes\cs{L}) \arrow{r} & (p^*\cs{L})\boxtimes (p^*\cs{L}) 
\end{tikzcd}
\]
Since this diagram commutes, and since the appropriate version of \eqref{eq:alphamu} also commutes, it follows that $\mu : m^*\cs{L} \to \cs{L}\boxtimes\cs{L}$ is a morphism of equivariant linear character sheaves.
The argument that $\vartheta : \cs{L}\boxtimes\cs{L} \to \theta^*(\cs{L}\boxtimes\cs{L})$ is also straightforward.
\end{proof}
